Supplement: Bootstrapping Domain Specific Languages for Bayesian Program Learning

### 2018-02-09 supplement.tex
initial DSL primitives and types

\section{Estimating $\theta$}

We write $c(e,p)$ to mean the number of times that primitive $e$ was used in program $p$;  $R(p)$ to mean the sequence of types input to sample in Alg.\ref{programGenerativeModel}. Jensen's inequality gives an intuitive lower bound on the likelihood of a program $p$:
\begin{align*}
  \log \probability[p|\theta]&\stackrel{+}{ = }\sum_{e\in \mathcal{D}} c(e,p)\log \theta_e - \sum_{\tau\in R(p)} \log \sum_{\substack{e:\tau'\in \mathcal{D}\\\text{unify}(\tau,\tau')}}\theta_e\\
  &\stackrel{+}{\geq }\sum_{e\in \mathcal{D}} c(e,p)\log \theta_e - c(p)\log \sum_{\tau\in R(p)} \sum_{\substack{e:\tau'\in \mathcal{D}\\\text{unify}(\tau,\tau')}}\theta_e\\
  & = \sum_{e\in \mathcal{D}} c(e,p)\log \theta_e - c(p)\log \sum_{e\in \mathcal{D}} r(e,p)\theta_e
\end{align*}
where $c(p) = \sum_{e\in \mathcal{D}}c(e,p)$ and $r(e:\tau',p) = \sum_{\tau\in R(p)} \indicator[\text{canUnify}(\tau,\tau')]$.

@@

>>>>

\section{DSL Induction}

\subsection{Structure learning}

We use Algorithm~\ref{grammarInductionAlgorithm} to search for the structure of the DSL that best explains the frontiers.

\begin{algorithm}[tb]
   \caption{DSL Induction Algorithm}
   \label{grammarInductionAlgorithm}
   \begin{algorithmic}
     \STATE {\bfseries Input:} Set of frontiers $\{\mathcal{F}_x\}$
     \STATE \textbf{Hyperparameters:} Pseudocounts $\alpha$, regularization parameter $\lambda$
     \STATE \textbf{Output:} DSL $\mathcal{D}$, weight vector $\theta$
     \STATE Define $L(\mathcal{D},\theta) =  \prod_x \sum_{z\in \mathcal{F}_x} \probability[z|\mathcal{D},\theta]$
     \STATE Define $\theta^*(\mathcal{D}) = \argmax_\theta \text{Dir}(\theta|\alpha) L(\mathcal{D},\theta)$
     \STATE Define $\text{score}(\mathcal{D}) = \log \probability[\mathcal{D}] + L(\mathcal{D},\theta^*) - ||\theta||_0$
     \STATE $\mathcal{D}\gets$ every primitive in $\{\mathcal{F}_x\}$
     \WHILE {true}
     \STATE N $\gets \{\mathcal{D}\cup \{s\} | x\in X, z\in \mathcal{F}_x, s\text{ subexpression of }z\}$
     \STATE $\mathcal{D}'\gets \argmax_{\mathcal{D}'\in N}\text{score}(\mathcal{D}') $
     \STATE \textbf{if }$\text{score}(\mathcal{D}') < \text{score}(\mathcal{D})$\textbf{ return }$\mathcal{D},\theta^*(\mathcal{D})$
     \STATE $\mathcal{D}\gets\mathcal{D}'$
     \ENDWHILE
   \end{algorithmic}
\end{algorithm}

\subsection{Estimating $\theta$}

We write $c(e,p)$ to mean the number of times that primitive $e$ was used in program $p$;  $R(p)$ to mean the sequence of types input to sample in Alg.~1 of the main paper. Jensen's inequality gives an intuitive lower bound on the likelihood of a program $p$:
\begin{align*}
  \log \probability[p|\theta]&\stackrel{+}{ = }\sum_{e\in \mathcal{D}} c(e,p)\log \theta_e - \sum_{\tau\in R(p)} \log \sum_{\substack{e:\tau'\in \mathcal{D}\\\text{unify}(\tau,\tau')}}\theta_e\\
  &\stackrel{+}{\geq }\sum_{e\in \mathcal{D}} c(e,p)\log \theta_e - c(p)\log \sum_{\tau\in R(p)} \sum_{\substack{e:\tau'\in \mathcal{D}\\\text{unify}(\tau,\tau')}}\theta_e\\
  & = \sum_{e\in \mathcal{D}} c(e,p)\log \theta_e - c(p)\log \sum_{e\in \mathcal{D}} r(e,p)\theta_e
\end{align*}
where $c(p) = \sum_{e\in \mathcal{D}}c(e,p)$ and $r(e:\tau',p) = \sum_{\tau\in R(p)} \indicator[\text{canUnify}(\tau,\tau')]$.

@@

\subsection{Boolean circuits}
\begin{center}
\tt
\begin{tabular}{| l | l |}
  \hline
  \textrm{\emph{name}} & \textrm{\emph{type}} \\
  \hline
    nand & $\text{bool}\to \text{bool}\to \text{bool}$ \\  \hline
\end{tabular}
\end{center}

\subsection{Symbolic regression}
\begin{center}
\tt
\begin{tabular}{| l | l |}
  \hline
  \textrm{\emph{name}} & \textrm{\emph{type}} \\
  \hline
  + & $\mathbb{R}\to \mathbb{R}\to \mathbb{R}$ \\
  *&  $\mathbb{R}\to \mathbb{R}\to \mathbb{R}$\\
  real&$\mathbb{R}$\\\hline
\end{tabular}
\end{center}

\subsection{String editing}
\begin{center}
\tt
\begin{tabular}{| l | l |}
  \hline
  \textrm{\emph{name}} & \textrm{\emph{type}} \\
  \hline
  0&$\mathbb{Z}$\\
  +1&$\mathbb{Z}\to \mathbb{Z}$\\
  -1&$\mathbb{Z}\to \mathbb{Z}$\\
  split&$\text{char}\to \text{string}\to \left[\text{string} \right]$\\
  join&$\text{string}\to \left[\text{string} \right]\to \text{string}$\\
  map&$(\text{string}\to \text{string})\to \left[\text{string} \right]\to \left[\text{string} \right]$\\
  nth&$\mathbb{Z}\to \left[\text{string} \right]\to \text{string}$\\
  slice&$\mathbb{Z}\to \mathbb{Z}\to \text{string}\to \text{string}$\\
  length&$\text{string}\to \mathbb{Z}$\\
  chr->str&$\text{char}\to \text{string}$\\
  trim&$\text{string}\to \text{string}$\\
  upper&$\text{string}\to \text{string}$\\
  lower&$\text{string}\to \text{string}$\\
  capitalize&$\text{string}\to \text{string}$\\\hline
\end{tabular}
\end{center}
We also provide the constant empty string which has type string and also provide
constant characters which have the type char.
